---
title: "p8105_hw1_qg2155"
author: "Helen Guan"
date: "9/18/2018"
output: github_document
---

Question #1
------------
 
```{r dataframe1}
library(tidyverse)

## set seed to get the same result every time for vec_random variable
set.seed(55)

## creating the dataframe
question_1 = tibble(
  vec_random = runif(10, 0, 5),
  vec_logical = vec_random > 2,
  vec_char = c("This", "is", "my", "answer", "to", "homework", "one", "question", "number", "one"),
  vec_factor = factor(c("yes", "no", "yes", "no", "yes", "no", "yes", "no", "yes", "no"))
)

## viewing the dataframe
question_1

## Taking the average of each variable
mean(question_1$vec_random)
mean(question_1$vec_logical)
mean(question_1$vec_char)
mean(question_1$vec_factor)
```
When trying to take the average of each variable, the output for variable vec_random is 1.620843 and vec_logical is 0.4. The results for vec_char and vec_factor are both NA. The error message for the character and factor vectors are that argument is not numeric or logical.  

```{r convert}
## transforming variable types
as.numeric(question_1$vec_logical)
as.numeric(question_1$vec_char)
as.numeric(question_1$vec_factor)
```
The logical and factor vectors were converted to numeric vector. But the character vector received error message stating that NAs introduced by coercion. The mean of the logical vector was 0.4 and fator vector was 1.5, but the mean of character variable still produced the output of NA.

```{r convert2}
## transforming character variable to factor to numeric and factor variable to character to numeric 
as.numeric(as.factor(question_1$vec_char))

as.numeric(as.character(question_1$vec_factor))
```
I was successfully able to convert the character variable to factor to numeric, and got a mean of 5.1. However, when I converted the factor variable to character to numeric, I got the same warning message that NAs were introduced by coercion. When I tried to take the mean of the new variable, the output was NA. It appears that character variables cannot be converted directly to numeric. Character variables first need to be converted to factor and then to numeric in order for it to work. 

Question #2
--------
```{r dataframe2}
## set seed to ensure reproducibility
set.seed(808)

## creating the dataframe
question_2 = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  xyvec_logic = x + y > 0,
  xyvec_num = as.numeric(xyvec_logic),
  xyvec_fac = as.factor(xyvec_logic)
)

## viewing the dataframe
question_2
```
This dataset has `r nrow(question_2)` rows and `r ncol(question_2)` columns. X and Y are random sample of size 1000 from a standard Normal distribution. xyvec_logic is a logical vector that indicates whether x + y is greater than 0. xyvec_num is a numeric and xyvec_fac is a factor conversion of the logical vector. The mean of x is `r mean(question_2$x)` and the median of x is `r median(question_2$x)`. The proportion of cases for which the logical vector is true is `r mean(question_2$xyvec_logic == TRUE)`.

Without true, it is `r mean(question_2$xyvec_logic)`.

```{r plots}
## creating scatterplot of y vs x using logical vector for color points
ggplot(question_2, aes(x = x, y = y, color = xyvec_logic)) + geom_point()
ggsave("scatter_plot.pdf", height = 4, width = 6)

## creating scatterplots y vs x using the numeric and factor vectors for color points
ggplot(question_2, aes(x = x, y = y, color = xyvec_num)) + geom_point()
ggplot(question_2, aes(x = x, y = y, color = xyvec_fac)) + geom_point()
```

